---
title: "Figure2_VAF_SBSsignature"
author: "Sanghoon Lee"
format: html
editor: visual
---

# Figure 2A and B. Variant allele frequency (VAF)

A. VAF of 11 WM mutated genes based on the proportion of variant DNA reads within the total number of reads at that locus.

B. Boxplot of the MYD88L265P VAF in different WM subtypes with no significant difference between MBC-like and PC-like.


## Lectures or useful websites

[Lecutre for VAF](https://bioconductor.org/packages/devel/bioc/vignettes/maftools/inst/doc/oncoplots.html)

## Step1. Load necessary R packages

```{r}
library(maftools) # bioconductor for oncoplots
library(data.table)
library(tidyverse)
```

## Step2. Set working directory and input file names

```{r}
dir <- dirname(rstudioapi::getSourceEditorContext()$path); print(dir) # 
setwd(dir); # "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N05b_WaldenstromMacroglobulinemia_Dylan/12_Github_WMWGS_Quarto/Figure2_VAF_SBSsignature"

MAFData_SNP_OutFile_NoFakeGene <- "WMWGS_Consensus_MAF_SNP_47s_ReadMAFMutInput_NoFakeGene.txt"
ClinAnnotFile_Indel <-  "WMWGS_Consensus_MAFSomaticIndel_47s_ClinicalAnnotInput.txt"

WMSubtypeData_RmvLowQualityTumorFile <- "WMSubtypeData_RmvLowQualityTumor.rds"
```

## Step3. Create VAF box plot (Figure 2A)

```{r}
laml_IndelSNV_HighFreqGene_NoFake = maftools::read.maf(maf=MAFData_SNP_OutFile_NoFakeGene, clinicalData=ClinAnnotFile_Indel) 
par(mfrow=c(1,1))
plotVaf(maf=laml_IndelSNV_HighFreqGene_NoFake, vafCol="t_vaf", top=20, width=4.5, height=3.7, fn="Fig2A_VAF_WM11MarkerGene_v1.0.pdf")

```

## Step4. Create MYD88 VAF box plot by different WM subtype (Figure 2B)

```{r}
MAFData_All47s_IndelSNV_TopGene_NoFake <- fread(MAFData_SNP_OutFile_NoFakeGene, header=TRUE, stringsAsFactors=FALSE) 
## Subset MYD88 t_vaf column.
MAFData_All47s_IndelSNV_MYD88 <- MAFData_All47s_IndelSNV_TopGene_NoFake %>% dplyr::filter(Hugo_Symbol=="MYD88") %>%
                        dplyr::select(Hugo_Symbol,Tumor_Sample_Barcode, t_vaf ); dim(MAFData_All47s_IndelSNV_MYD88) # 44 3


WMSubtypeData_RmvLowQualityTumor <- readRDS(WMSubtypeData_RmvLowQualityTumorFile)

MAFData_All47s_IndelSNV_MYD88_Subtype <- dplyr::inner_join(MAFData_All47s_IndelSNV_MYD88,
                              WMSubtypeData_RmvLowQualityTumor[,c("Tumor_Sample_Barcode","Diagnosis","WMSubtype")] );


###### Boxplot
my_comparisons <- list( c("G1_MBC_LIKE", "G2_PC_LIKE"), c("G2_PC_LIKE", "G3_NotAvail"), c("G1_MBC_LIKE", "G3_NotAvail") )
MyTitle <- "MYD88 VAF by different WM subtype"

MyBoxplot <- MAFData_All47s_IndelSNV_MYD88_Subtype %>% ggplot(aes(x=WMSubtype,y=t_vaf, fill=WMSubtype)) +
  geom_boxplot() + viridis::scale_fill_viridis(discrete = TRUE, alpha=0.)  +
  geom_jitter(aes(colour=Diagnosis),  size=2, alpha=0.9, width=0.2) +
  scale_color_manual(values = c("ND" = "blue", "Relapse" = "red", "NotAvail"="grey")) +
  ylim(floor(0), 1.5) + # hrbrthemes::theme_ipsum()  # This causes error
  theme( plot.title = element_text(size=11)) + theme_bw() + # legend.position="none",
  ggtitle(MyTitle) + xlab("")  + labs(y="Variant allel frequency (VAF)")  # +  theme(text = element_text(family = "Arial", size = 12)) 
MyBoxplot <- MyBoxplot+ scale_fill_manual(values=c("plum1", "gold2","lightblue1")) + # theme(legend.position="none")
  ggpubr::stat_compare_means(method="wilcox.test", comparisons = my_comparisons) + # Add pairwise comparisons p-value()  # method="t.test"
  ggpubr::stat_compare_means(label.y = 15) +    # Add global p-value
  theme(axis.line = element_line(colour = "black"), ## remove grid inside ggplot
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

MyBoxplot

OutFileName<-"Fig2B_MYD88_VAF_ByWMSubtype_v1.0.pdf"
ggsave(file=OutFileName, MyBoxplot, width=4.3, height=3)  ## save plot

```




# Figure 2C. Single base Substitution (SBS) signature 

C. SBS signatures by WM subtype showing relative contribution of SBS9, SBS1, SBS5, SBS8, SBS18, and SBS48.

## Lectures or useful websites

[Mutational signature analysis](https://github.com/pblaney/mmsig)

## Step1. Load necessary R packages

```{r}
## To use hg38 aligned SNV calling data and COSMIC v3.1 mutational signatures. I should install mmsig R package.
library(deconstructSigs)
library(mmsig)
library(ggplot2)
library(scales) ## To control decimal places in y-axis in ggplot.   label_number()
library(ggpubr) ## to use stat_compare_means() that calculates wilcoxon test

```

## Step2. Set input file names

```{r}
MAFDatahg38File <- "MAFData_All47s_SNP.rds"
WMSubtype47sFile <- "WMSubtypeData_RmvLowQualityTumor.rds"

```



## Step3. Load hg38 data of SNV MAF file.

```{r}
MAFDatahg38 <- readRDS(MAFDatahg38File); dim(MAFDatahg38); MAFDatahg38[1:2,] #  116891    114 from folder 06b  # 116895 15 after Spotcheck and adding MYD88 SNV 4 smp
#       Hugo_Symbol Entrez_Gene_Id Center NCBI_Build Chromosome Start_Position End_Position Strand Variant_Classification Variant_Type Reference_Allele Tumor_Seq_Allele1 Tumor_Seq_Allele2     dbSNP_RS
# 1       PRKCZ              0      .     GRCh38       chr1        2141349      2141349      +                 Intron          SNP                T                 T                 C rs1039845054
length(unique(MAFDatahg38$Tumor_Sample_Barcode)) ## 47


MAFDatahg38_Subset <- MAFDatahg38 %>% dplyr::select(Tumor_Sample_Barcode, Chromosome, Start_Position,Reference_Allele, Tumor_Seq_Allele2 )
colnames(MAFDatahg38_Subset) <- c("sample","chr","pos","ref","alt"); MAFDatahg38_Subset[1:2,]
#   sample  chr     pos ref alt
# 1 01_076 chr1 2141349   T   C
# 2 01_076 chr1 3079529   C   T
table(MAFDatahg38_Subset$sample)[1:3]
# 01_076 01_115 01_131
# 2377   3293   1897

```

## Step4. Subset Signature Reference

```{r}
# remove the platinum signature (SBS35) because the patients are not platinum exposed
data(signature_ref_cosmic_v3_1_hg38)
sig_ref_cosmic_v3_1_hg38 <- signature_ref_cosmic_v3_1_hg38[c("sub", "tri", "SBS1", "SBS2", "SBS5", "SBS8", "SBS9", "SBS13", "SBS18", "SBS84")]
# Mutation data is already subset to same 3 samples as hg19 example

dim(sig_ref_cosmic_v3_1_hg38); sig_ref_cosmic_v3_1_hg38[1,] #  96 10
#   sub tri         SBS1         SBS2       SBS5       SBS8        SBS9       SBS13      SBS18  SBS84
# 1 C>A ACA 0.0008760229 5.790059e-07 0.01205205 0.04431064 0.000560639 0.001816879 0.05168832 0.00350024


### Perform Mutational Signature Analysis
# Bootstrapping large datasets with many iterations can significantly increase runtime.
set.seed(1)
sig_out_cosmic_v3_1_hg38 <- mm_fit_signatures(muts.input = MAFDatahg38_Subset,
                                              sig.input = sig_ref_cosmic_v3_1_hg38,
                                              input.format = "vcf",
                                              sample.sigt.profs = NULL,
                                              strandbias = TRUE,
                                              bootstrap = TRUE,
                                              iterations = 200,
                                              genome = "hg38",
                                              refcheck = TRUE,
                                              cos_sim_threshold = 0.01,
                                              force_include = c("SBS1", "SBS5"),
                                              dbg = FALSE)
class(sig_out_cosmic_v3_1_hg38) # list
saveRDS(sig_out_cosmic_v3_1_hg38, "sig_out_cosmic_v3_1_hg38.rds")

####### Sort samples by SBS9 proportion
MutSigEst <- sig_out_cosmic_v3_1_hg38$estimate; class(MutSigEst); dim(MutSigEst) # data.frame # 47 9
MutSigEst_SortBySBS9 <- MutSigEst[with(MutSigEst, order(MutSigEst$SBS9, decreasing=FALSE)),]
SampleNameSortedBySBS9 <- rownames(MutSigEst_SortBySBS9)

```

## Step 5. plot_signatures()

```{r}
plot_signatures <- function (sig, sig_order = c("SBS1", "SBS2", "SBS13", "SBS5",  "SBS8", "SBS9", "SBS18", "SBS-MM1", "SBS35"), samples = FALSE) {
  rotatedAxisElementText = function(angle, position = "x") {
    angle = angle[1]
    position = position[1]
    positions = list(x = 0, y = 90, top = 180, right = 270)
    if (!position %in% names(positions))
      stop(sprintf("'position' must be one of [%s]", paste(names(positions),  collapse = ", ")), call. = FALSE)
    if (!is.numeric(angle))
      stop("'angle' must be numeric", call. = FALSE)
    rads = (angle - positions[[position]]) * pi/180
    hjust = 0.5 * (1 - sin(rads))
    vjust = 0.5 * (1 + cos(rads))
    element_text(angle = angle, vjust = vjust, hjust = hjust)
  }
  scale_fill_sigs <- function(...) {
    ggplot2:::manual_scale("fill", values = setNames(c(RColorBrewer::brewer.pal(8,"Dark2"), "lightblue", "#FB8072"),
                            c("SBS1", "SBS2", "SBS5", "SBS8", "SBS9", "SBS13", "SBS18", "SBS-MM1", "SBS35", "SBS84")), ...)
  }
  sig_Proc <- sig %>% rownames_to_column(var = "sample") %>%  dplyr::select(-mutations) %>% reshape2::melt(id.var = "sample",
                                       variable.name = "SBS", value.name = "prop") %>% mutate(SBS = factor(SBS, levels = sig_order))
  sig_Proc$sample <- factor(sig_Proc$sample, levels = SampleNameSortedBySBS9)

  sigPlot <- sig_Proc %>% ggplot(aes(sample, prop, fill = SBS)) +
                      geom_col(width = 1) + scale_fill_sigs() + scale_y_continuous(expand = c(0, 0)) + theme_bw() + labs(x = "Sample", y = "Relative contribution",
                     fill = "Signature") + theme(text = element_text(size = 10, color = "black"), axis.text = element_text(color = "black"),
                     axis.text.y = element_text(size = 12), axis.title = element_text(size = 14), panel.grid = element_blank(), panel.border = element_blank(),
                     legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 14), axis.title.x = element_blank())
  if (samples) {
    sigPlot + theme(axis.text.x = rotatedAxisElementText(90,  "top"))
  }
  else {
    sigPlot + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
}

```

## Step6. Plot Signature Estimates  - all 47 samples.

```{r}
### Find SBS that don't hava any value across 47 s
colnames(MutSigEst_SortBySBS9)[colSums((MutSigEst_SortBySBS9))==0] # [1] "SBS2"  "SBS13"
## Remove SBS2 and SBS13 columns.
MutSigEst_SortBySBS9_RmvSBS2SBS13 <- MutSigEst_SortBySBS9 %>% dplyr::select(-c("SBS2","SBS13"))


MyPlotSignature <- plot_signatures(sig = MutSigEst_SortBySBS9_RmvSBS2SBS13, samples = T,
                                   sig_order = c("SBS9","SBS1", "SBS5", "SBS8", "SBS18", "SBS84"))

SigPlotOutName <- "Out_MutSigBarplot_WM_SNV_47s_AfterSpotcheck_SortedBySBS9Increasing.pdf"  # Line 80 determine decreasing or increasing.
ggsave(filename=SigPlotOutName, MyPlotSignature, width=9, height=5)

```
## Step7. Plot Signature Estimates - split by WM subtypes.

```{r} 
## Read WM Subtype file.
WMSubtype47s <- readRDS(WMSubtype47sFile); WMSubtype47s[1:2,]
#     Tumor_Sample_Barcode Diagnosis Treatment   WMSubtype
# 1                 01_076        ND Rituximab G1_MBC_LIKE

## Split mutation signature data by WM subtype
## 1) G1_MBC_Like
MutSigEst_SortBySBS9_MBC <- MutSigEst_SortBySBS9_RmvSBS2SBS13 %>% dplyr::filter(rownames(MutSigEst_SortBySBS9_RmvSBS2SBS13) %in% WMSubtype47s$Tumor_Sample_Barcode[WMSubtype47s$WMSubtype=="G1_MBC_LIKE"])
dim(MutSigEst_SortBySBS9_MBC) # 10 9.

MyPlotSignature_MBC <- plot_signatures(sig = MutSigEst_SortBySBS9_MBC, samples = T,
                                   sig_order = c("SBS9","SBS1", "SBS5", "SBS8", "SBS18", "SBS84"))

SigPlotOutName <- "Fig2C_MutSigBarplot_WM_SNV_MBCLike10s_AfterSpotcheck_SortedBySBS9Increasing.pdf"  # Line 80 determine decreasing or increasing.
ggsave(filename=SigPlotOutName, MyPlotSignature_MBC, width=4, height=5)


## 2) G2_PC_Like
MutSigEst_SortBySBS9_PC <- MutSigEst_SortBySBS9_RmvSBS2SBS13 %>% dplyr::filter(rownames(MutSigEst_SortBySBS9_RmvSBS2SBS13) %in% WMSubtype47s$Tumor_Sample_Barcode[WMSubtype47s$WMSubtype=="G2_PC_LIKE"])
dim(MutSigEst_SortBySBS9_PC) # 22 9.

MyPlotSignature_PC <- plot_signatures(sig = MutSigEst_SortBySBS9_PC, samples = T,
                                      sig_order = c("SBS9","SBS1", "SBS5", "SBS8", "SBS18", "SBS84"))

SigPlotOutName <- "Fig2C_MutSigBarplot_WM_SNV_PCLike22s_AfterSpotcheck_SortedBySBS9Increasing.pdf"  # Line 80 determine decreasing or increasing.
ggsave(filename=SigPlotOutName, MyPlotSignature_PC, width=5.5, height=5)


## 3) G3_NotAvail
MutSigEst_SortBySBS9_NotAvail <- MutSigEst_SortBySBS9_RmvSBS2SBS13 %>% dplyr::filter(rownames(MutSigEst_SortBySBS9_RmvSBS2SBS13) %in% WMSubtype47s$Tumor_Sample_Barcode[WMSubtype47s$WMSubtype=="G3_NotAvail"])
dim(MutSigEst_SortBySBS9_NotAvail) # 15 9

MyPlotSignature_NotAvail <- plot_signatures(sig = MutSigEst_SortBySBS9_NotAvail, samples = T,
                                            sig_order = c("SBS9","SBS1", "SBS5", "SBS8", "SBS18", "SBS84"))

SigPlotOutName <- "Fig2C_MutSigBarplot_WM_SNV_NotAvailLike15s_AfterSpotcheck_SortedBySBS9Increasing.pdf"  # Line 80 determine decreasing or increasing.
ggsave(filename=SigPlotOutName, MyPlotSignature_NotAvail, width=4.5, height=5)


```







