---
title: "Figure3_CopyNumberAbnormality"
author: "Sanghoon Lee"
format: html
editor: visual
---

## Figure 3. Copy number abnormality profiles

**A.** Cumulative CNA across all chromosomes in 47 WM cases.

**B.** Expanded view of chromosome 1q.

**C.** Expanded view of chromosome 6q in 24 WM cases.

**D.** Expanded view of chromosome 6q across WM subtypes and an ideogram for B cell–associated genes in the region.

## Lectures or useful websites

[Lecutre for GenVisR](https://genviz.org/module-03-genvisr/0003/05/01/cnFreq_GenVisR/)

## Step1. Load necessary R packages

```{r}
library(karyoploteR)
library(GenVisR) # Bioconductor
library(dplyr)
library(GenomicRanges)
library(ggplot2) # facet_grid()
library(circlize) # rand_color(n=4)  To choose random colors.
library(data.table)

dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir); print(dir) #  "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N05b_WaldenstromMacroglobulinemia_Dylan/12_Github_WMWGS_Quarto/Figure3_CopyNumberAbnormality"

### Metadata
MetadataProcFile <- "Metadata_Proc.rds"
Chr6GeneAnnotFile <- "GCF_000001405.40_GRCh38.p14_genomic_Chr6.txt"     ## You can make this file yourself by extracting only Chr6 region. Otherwise, email me. 
source("Source_GenVisR_Modified_ByCNVSegment_BycnQual.R")
ChromosomeArmBandStartStopFile <-  "ChromosomeArmBandStartStop.rds"
CentromereDataFile <- "Centromere_1bpPlaceholder_UCSC_hg38.txt"
ChromosomeArmSubtypeCNVStateFile <- "ChromosomeArmSubtypeCNVState_AllSmp47_ManualProc_v1.2.txt"
```

## Step2. Read input data.

```{r}
cnData_NoWM91WM93WM94 <- readRDS("cnData_NoWM91WM93WM94.rds")

## metadata for MBC-like and PC-like information 
MetadataProc <- readRDS(MetadataProcFile); dim(MetadataProc); MetadataProc[1:2,]
colnames(MetadataProc)[colnames(MetadataProc)=="SmpName"] <- "sample"

## inner_join between MetadataProc and cnData_NoWM91WM93WM94. 
cnData_NoWM91WM93WM94_Metadata <- dplyr::inner_join(cnData_NoWM91WM93WM94, MetadataProc[, c("sample","Subtype")] ); dim(cnData_NoWM91WM93WM94_Metadata) # 2189 18 
# saveRDS(cnData_NoWM91WM93WM94_Metadata, file="cnData_NoWM91WM93WM94_Metadata.rds")

## split to MBC-like, PC-like or NotAvail 
cnData_NoWM91WM93WM94_MBCLike <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(Subtype=="G1_MBC_LIKE") %>% select(-Subtype) ; dim(cnData_NoWM91WM93WM94_MBCLike) # 493 17
cnData_NoWM91WM93WM94_PCLike <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(Subtype=="G2_PC_LIKE") %>% select(-Subtype) ; dim(cnData_NoWM91WM93WM94_PCLike) # 964 17
cnData_NoWM91WM93WM94_NotAvail <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(Subtype=="G3_NotAvail") %>% select(-Subtype) ; dim(cnData_NoWM91WM93WM94_NotAvail) # 693 17
```

## Step3. Create Figure 3A

```{r}
# Run it
# pdf("/Users/gagled01/morganLab/Waldenstroms/WGS/figures/GenVisR_Overall_CNVPlot_Draft.pdf", height = 8, width = 20)
# pdf("GenVisR_Overall_CNVPlot_Draft.pdf", height = 8, width = 20)
#       cnFreq(cnData, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free")
# dev.off()

## CentromereRegion <- ggplot2::geom_vline(xintercept = c(123400000,123400001,   93900000,93900001,    90900000, 90900001), linetype = "dashed", size=0.2)  # This doesn't work. 

chr_levels <- paste0("chr", c(1:22, "X", "Y"))
cnData_NoWM91WM93WM94$chromosome <- factor(cnData_NoWM91WM93WM94$chromosome, levels = chr_levels)

# centromere_df <- data.frame(chromosome = c("chr1", "chr2", "chr3"), centromere = c(123400000, 93900000, 90900000) )
# centromere_df$chromosome <- factor(centromere_df$chromosome, levels = chr_levels)

## Read Centromere 1bp placeholder data
centromere_df <- fread(CentromereDataFile, header=TRUE, stringsAsFactors=FALSE); dim(centromere_df); centromere_df[1:2,] # 24 3
#       Chromosome     Start       End
#        <char>     <int>     <int>
# 1:       chr1 123400000 123400001
colnames(centromere_df)[1:2]<- c('chromosome', 'centromere')
centromere_df$centromere[centromere_df$chromosome=="chr22"]<- 15000000
centromere_df$chromosome <- factor(centromere_df$chromosome, levels = chr_levels)

CentromereRegion <- ggplot2::geom_vline(data = centromere_df, aes(xintercept = centromere), linetype = "dashed", linewidth = 0.2)


## 1) All chromosome all samples
# CNVPlot_NoWM91WM93WM94<- GenVisR::cnFreq(cnData_NoWM91WM93WM94, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free") # This is original working. 
CNVPlot_NoWM91WM93WM94_Centromere <- GenVisR::cnFreq(cnData_NoWM91WM93WM94, genome="hg38", plotLayer=CentromereRegion) + ggplot2::facet_grid(. ~ chromosome, scales="free") ## If I want to draw vertical lines of the centromere region in the GenVisR plot. 
# print(CNVPlot_NoWM91WM93WM94_Centromere)
ggplot2::ggsave(filename="Fig3A_GenVisR_Overall_CNVPlot_NoWM91WM93WM94_AllSubject_Centromere.pdf", CNVPlot_NoWM91WM93WM94_Centromere, width=20, height=8)


# CentromereRegion <- ggplot2::geom_vline(xintercept = c(59800000, 59800001))
# ggplot2::ggsave(filename="GenVisR_Overall_CNVPlot_NoWM91WM93WM94_AllSubject_v1.3.pdf", CNVPlot_NoWM91WM93WM94,plotLayer=CentromereRegion, width=20, height=8)



## 2) All chromosome, MBC-Like 
length(unique(cnData_NoWM91WM93WM94_MBCLike$sample)) # 10
CNVPlot_NoWM91WM93WM94_MBC <- GenVisR::cnFreq(cnData_NoWM91WM93WM94_MBCLike, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free")
ggplot2::ggsave(filename="GenVisR_Overall_CNVPlot_AllChromosome_MBCLike10s_v1.1.pdf", CNVPlot_NoWM91WM93WM94_MBC, width=20, height=8)

## 3) All chromosome, PC-Like 
length(unique(cnData_NoWM91WM93WM94_PCLike$sample)) # 22
CNVPlot_NoWM91WM93WM94_PC <- GenVisR::cnFreq(cnData_NoWM91WM93WM94_PCLike, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free")
ggplot2::ggsave(filename="GenVisR_Overall_CNVPlot_AllChromosome_PCLike22s_v1.0.pdf", CNVPlot_NoWM91WM93WM94_PC, width=20, height=8)

## 4) All chromosome, NotAvail
length(unique(cnData_NoWM91WM93WM94_NotAvail$sample)) # 15
CNVPlot_NoWM91WM93WM94_NotAvail<- GenVisR::cnFreq(cnData_NoWM91WM93WM94_NotAvail, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free")
ggplot2::ggsave(filename="GenVisR_Overall_CNVPlot_AllChromosome_NotAvail15s_v1.0.pdf", CNVPlot_NoWM91WM93WM94_NotAvail, width=20, height=8)


# Just chr6
# blimp1 <- ggplot2::geom_vline(xintercept = c(82169987, 89926528,  105992690, 142751469)) # IBTK, BACH2, BLIMP1, HIVEP2
# pdf("/Users/gagled01/morganLab/Waldenstroms/WGS/figures/GenVisR_Chr6_CNVPlot.pdf", height = 8, width = 20)
# cnFreq(cnData, genome="hg38", plotChr = "chr6", plotLayer = blimp1)
# dev.off()

CentromereRegion <- ggplot2::geom_vline(xintercept = c(59800000, 59800001))  # Centromere region
CNVPlot_Chr6 <- cnFreq(cnData_NoWM91WM93WM94, genome="hg38", plotChr = "chr6", plotLayer = CentromereRegion); 
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr6_v1.2.pdf", CNVPlot_Chr6, width=5, height=4)



#cnSpec(cnData, genome="hg38")
```

## Step4. Figure 3B-C: Make CNV plot for individual subject, only Chr6q or only Chr1q

```{r}
dim(cnData_NoWM91WM93WM94_Metadata); table(cnData_NoWM91WM93WM94_Metadata$sample) # 2142  18
# 01-076 01-115 01-131 01-163 01-190 04-006   WM22   WM26   WM27   WM29   WM30   WM31   WM32   WM33   WM37   WM42   WM43   WM46   WM47   WM48   WM49   WM52   WM54   WM65   WM66   WM67   WM68   WM69   WM70   WM72 
#     36     43     40     40     88     40     41     44     37     44     47     46     44     62     37     48     72     41     47     39     42     33     30     37     43     32     50     44     59     37 
#   WM73   WM74   WM77    WM8   WM80   WM82   WM83   WM84   WM85   WM86   WM87   WM89    WM9   WM90   WM95   WM96   WM97 
#     45     35     36     57     43     51     49     99     40     43     43     37     36     71     53     31     30 

AllSubject <- unique(cnData_NoWM91WM93WM94_Metadata$sample); print(AllSubject) 
#  [1] "01-076" "01-115" "01-131" "01-163" "01-190" "04-006" "WM22"   "WM26"   "WM27"   "WM29"   "WM30"   "WM31"   "WM32"   "WM33"  
# [15] "WM37"   "WM42"   "WM43"   "WM46"   "WM47"   "WM48"   "WM49"   "WM52"   "WM54"   "WM65"   "WM66"   "WM67"   "WM68"   "WM69"  
# [29] "WM70"   "WM72"   "WM73"   "WM74"   "WM77"   "WM8"    "WM80"   "WM82"   "WM83"   "WM84"   "WM85"   "WM86"   "WM87"   "WM89"  
# [43] "WM9"    "WM90"   "WM95"   "WM96"   "WM97"  

for (EachSubject in AllSubject) {
    ## EachSubject <- AllSubject[10]; print(EachSubject)
    ## subset by each subject 
    cnData_NoWM91WM93WM94_EachSubject <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(sample %in% EachSubject) %>% select(-Subtype); dim(cnData_NoWM91WM93WM94_EachSubject) # 36 17

    CNVPlot_NoWM91WM93WM94_EachSubject <- GenVisR::cnFreq(cnData_NoWM91WM93WM94_EachSubject, genome="hg38") + ggplot2::facet_grid(. ~ chromosome, scales="free")
    CNVPlot_NoWM91WM93WM94_EachSubject  ## Plot display doesn't work 
    ggplot2::ggsave(filename=paste0("GenVisR_Overall_CNVPlot_NoWM91WM93WM94_",EachSubject ,"_v1.2.pdf"), CNVPlot_NoWM91WM93WM94_EachSubject, width=20, height=8)
} 

```

## Step5. Make a table to summarize Gain or Loss in each chromosome p/q arm and in each sample

```{r}
ChromosomeArmBandStartStop<- readRDS(ChromosomeArmBandStartStopFile); dim(ChromosomeArmBandStartStop); ChromosomeArmBandStartStop[1:2,] # 862 5
#   Chromosome Arm   Band    Start    Stop
#   <chr>      <chr> <chr>   <int>   <int>
# 1 chr1       p     36.33       1 2300000
# 2 chr1       p     36.32 2300001 5300000

##### I should add 0 in front of chr"1" ... chr"9" to sort Chr#_Gain|Loss by the order of Chromosome numbers
ChromosomeArmBandStartStop$Chromosome <- ifelse(grepl("chr\\d$", ChromosomeArmBandStartStop$Chromosome), gsub("chr","chr0",ChromosomeArmBandStartStop$Chromosome ), 
                                            ChromosomeArmBandStartStop$Chromosome)

########### filter only CNV state =1 or > 3
cnData_GainLoss_AllSmp <- data.frame()
for (EachSubject in AllSubject) {
    ## EachSubject <- AllSubject[2]; print(EachSubject)
    ## subset by each subject 
    cnData_NoWM91WM93WM94_EachSubject <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(sample %in% EachSubject);     
    dim(cnData_NoWM91WM93WM94_EachSubject) # 36 17
    ## filter the bins that have gain >= 3 or loss==1, and "end-start > 1e+6
    cnData_GainLoss <- cnData_NoWM91WM93WM94_EachSubject %>% data.frame %>% dplyr::filter((segmean==1|segmean >=3)&((end-start)>=1e+6 ) ) %>%
                                      dplyr::filter(!chromosome %in% c("chrM","chrX","chrY","chr23")) %>% dplyr::select(c(chromosome, start, end,segmean, sample, Subtype));
    dim(cnData_GainLoss)
    if(nrow(cnData_GainLoss)==0) {
          cnData_GainLoss <- data.frame("chromosome"="NotAvail", "start"=0, "end"=0,"segmean"=0, 
                                        "sample"=EachSubject, "Subtype"=unique(cnData_NoWM91WM93WM94_EachSubject$Subtype))
    }
    cnData_GainLoss_AllSmp <- dplyr::bind_rows(cnData_GainLoss_AllSmp,cnData_GainLoss )
    
    ##### I should add 0 in front of chr"1" ... chr"9" to sort Chr#_Gain|Loss by the order of Chromosome numbers
    cnData_GainLoss_AllSmp$chromosome <- ifelse(grepl("chr\\d$", cnData_GainLoss_AllSmp$chromosome), gsub("chr","chr0",cnData_GainLoss_AllSmp$chromosome ), cnData_GainLoss_AllSmp$chromosome)
    
}    ## end of for loop 
dim(cnData_GainLoss_AllSmp) # 296 6

########### ########################## Annotate the bin region with p or q arm 

RowData_BindRow <- data.frame(); LoopNumb<-0
TotalNumb_Bin <- nrow(cnData_GainLoss_AllSmp); print(TotalNumb_Bin) # 295
for (i in 1:TotalNumb_Bin ) {
    # i<-248
    LoopNumb<-LoopNumb+1; print(paste0("LoopNumb: ", LoopNumb))
    RowData <- cnData_GainLoss_AllSmp[i, ]
    
    if(RowData$start != 0) {  ## NotAvail has start position 0 
          ## Extract the Chromosome Position from "ChromosomeArmBandStartStop" by Chr#, start, end 
          ChromosomeArmBandStartStop_SubsetBystart <- ChromosomeArmBandStartStop %>% dplyr::filter(Chromosome==RowData$chromosome, Start <= RowData$start, Stop >= RowData$start) 
             # print(ChromosomeArmBandStartStop_SubsetBystart)
          ChromosomeArmBandStartStop_SubsetByend <- ChromosomeArmBandStartStop %>% dplyr::filter(Chromosome==RowData$chromosome, Start <= RowData$end, Stop >= RowData$end) 
            #  print(ChromosomeArmBandStartStop_SubsetByend)
        
          if(ChromosomeArmBandStartStop_SubsetBystart$Arm == ChromosomeArmBandStartStop_SubsetByend$Arm) {
                ArmPQ <- as.character(ChromosomeArmBandStartStop_SubsetBystart$Arm)
          } else if (ChromosomeArmBandStartStop_SubsetBystart$Arm != ChromosomeArmBandStartStop_SubsetByend$Arm) {
                  # ArmPQ <- "pqCentromere"
            
                  ### Some bins have ArmPQ: "pqCentromere" I should replace them with p, q, ChromosomeWhole
                  if(RowData$chromosome %in% c("chr02","chr11","chr21","chr22")) {
                         ArmPQ <- "p"
                  } else if (RowData$chromosome %in% c("chr01","chr08","chr09","chr10","chr13","chr14","chr15","chr17","chr20")) {
                         ArmPQ <- "q"
                  } else if (RowData$chromosome %in% c("chr03") & RowData$end == 198215400) {
                         ArmPQ <- "q"
                  } else if (RowData$chromosome %in% c("chr03", "chr04") & RowData$start <= 13000 & RowData$end >= 190000000) {
                         ArmPQ <- "ChromosomeWhole"
                  } else if (RowData$chromosome %in% c("chr04") & RowData$start == 49072176 ) {
                         ArmPQ <- "p"
                  } else if (RowData$chromosome %in% c("chr04") & RowData$start >= 35000000 ) {
                         ArmPQ <- "q"
                  } else if (RowData$chromosome %in% c("chr05") & RowData$start == 46427400 ) {
                         ArmPQ <- "p"
                  } else if (RowData$chromosome %in% c("chr06") & RowData$start <= 130000 & RowData$end >= 60000000) {
                         ArmPQ <- "p"
                  } else if (RowData$chromosome %in% c("chr06") & RowData$start == 39081429 ) {
                         ArmPQ <- "q"
                  } else if (RowData$chromosome %in% c("chr19") ) {
                         ArmPQ <- "q"
                  } else if (RowData$chromosome %in% c("chr12", "chr16","chr18") ) {
                         ArmPQ <- "ChromosomeWhole"
                  } else { 
                         ArmPQ <- "pqCentromere"
                  }
          }
          ArmBandStart <-  ChromosomeArmBandStartStop_SubsetBystart$Band
          ArmBandEnd <- ChromosomeArmBandStartStop_SubsetByend$Band
    } else {
          ArmPQ <- "NotAvail"
    }
    RowData$ChrArm <- ArmPQ
    RowData$ArmBandStart <- ArmBandStart
    RowData$ArmBandEnd <- ArmBandEnd
    #### Add one more column to indicate 'Chr#_Loss' or 'Chr#_Gain'  
    # RowData$ChrCNVState<-ifelse(RowData$segmean==1, paste0(RowData$chromosome,  "_Loss"), ifelse(RowData$segmean >= 3, paste0(RowData$chromosome,"_Gain"),
    #                               ifelse(RowData$segmean==0, "NotAvail","")))

    #### dplyr::bind_rows 
    RowData_BindRow <- dplyr::bind_rows(RowData_BindRow, RowData) 
     
}
dim(RowData_BindRow); RowData_BindRow[1:2,]; # 299   9
# RowData_BindRow_PQCentromere <- RowData_BindRow %>% dplyr::filter(ChrArm=="pqCentromere")
# RowData_BindRow_Chr1q <- RowData_BindRow %>% dplyr::filter(chromosome=="chr01", ChrArm=="q")

#### Add one more column to indicate 'Chr#_p' or 'Chr#_q'  
RowData_BindRow_PQArm <- RowData_BindRow %>% dplyr::mutate(ChrPQArm = paste0(chromosome, "_",ChrArm),  
                                   ChrCNVState=ifelse(RowData_BindRow$segmean==1,"Loss", ifelse(RowData_BindRow$segmean >= 3,"Gain",ifelse(RowData_BindRow$segmean==0, "NotAvail",""))))  # %>%
                                 #  dplyr::filter(!grepl("NotAvail",RowData_BindRow$chromosome)) 

table(RowData_BindRow_PQArm$ChrPQArm) ; dim(RowData_BindRow_PQArm) #  299  11
# chr01_Gain chr01_Loss chr02_Gain chr02_Loss chr03_Gain chr03_Loss chr04_Gain chr04_Loss chr05_Gain chr05_Loss chr06_Gain chr06_Loss chr07_Gain chr07_Loss chr08_Gain chr08_Loss chr09_Gain 
#         17         16         15          1          6          3         12          4          3          1          9         25          3          1          7         13         13 
# chr09_Loss chr10_Gain chr10_Loss chr11_Gain chr11_Loss chr12_Gain chr12_Loss chr13_Gain chr13_Loss chr14_Gain chr14_Loss chr15_Gain chr15_Loss chr16_Gain chr17_Gain chr17_Loss chr18_Gain 
#          5          3          1          3          1          4          1         13          7          8          1         23          2          4          5          3          7 
# chr19_Gain chr20_Gain chr21_Gain chr22_Gain   NotAvail 
#          2         10         28         15          4 

#### I should find duplicated rows by 'sample', 'ChrPQArm', 'ChrCNVState' and remove them. 
RowData_BindRow_PQArm_sort <- RowData_BindRow_PQArm %>% arrange(chromosome, sample); 
print(RowData_BindRow_PQArm_sort[5:7,])
#   chromosome     start       end segmean sample     Subtype ChrArm ChrPQArm ChrCNVState
# 5      chr01  69702329  72828300       1 01-190 G1_MBC_LIKE      p  chr01_p        Loss
# 6      chr01  73170435  79058766       1 01-190 G1_MBC_LIKE      p  chr01_p        Loss
# 7      chr01 168079200 171494292       1 01-190 G1_MBC_LIKE      q  chr01_q        Loss

RowData_BindRow_PQArm_NoDup <- RowData_BindRow_PQArm_sort %>% dplyr::mutate(SmpChrPQArmCNVState=paste0(sample,"_",ChrPQArm,"_",ChrCNVState)) %>% 
          dplyr::filter(!duplicated(SmpChrPQArmCNVState)); dim(RowData_BindRow_PQArm_NoDup) # 266 10 
# View(RowData_BindRow_PQArm_NoDup)
length(unique(RowData_BindRow_PQArm_NoDup$sample)) # 47

################### Extract rows by sampleID and do bind_row() 
ChromosomeArmSubtypeCNVState_AllSmp <- data.frame(); ForLoopNumb<-0
for (EachSmpID in AllSubject) {
      # EachSmpID <- AllSubject[1]
      ForLoopNumb <- ForLoopNumb + 1; print(paste0("Current for loop numb:", ForLoopNumb))
      ChromosomeArmSubtypeCNVState_Subset <- RowData_BindRow_PQArm_NoDup %>% dplyr::filter(sample==EachSmpID) %>% dplyr::select(c(sample,Subtype, ChrPQArm, ChrCNVState )) #
                      # %>% tibble::column_to_rownames("ChrPQArm") %>% t %>% data.frame; print(ChromosomeArmSubtypeCNVState_Subset)
      ## fill the empty data.frame with the values I found.  
      
      ###################  Empty data.frame of the final table
      FinalDF_ChrPQArm_GainLoss <- data.frame(matrix(nrow=1, ncol=(length(unique(RowData_BindRow_PQArm_NoDup$ChrPQArm)))+2))  ### Total number of columns is 50
      colnames(FinalDF_ChrPQArm_GainLoss) <- c("SampleID","Subtype", sort(unique(RowData_BindRow_PQArm_NoDup$ChrPQArm)) )  ## ncol=50

      for(i in 1:nrow(ChromosomeArmSubtypeCNVState_Subset)) {
          # i<-1
          FinalDF_ChrPQArm_GainLoss$SampleID=ChromosomeArmSubtypeCNVState_Subset[i,]$sample
          FinalDF_ChrPQArm_GainLoss$Subtype=ChromosomeArmSubtypeCNVState_Subset[i,]$Subtype
          FinalDF_ChrPQArm_GainLoss[,colnames(FinalDF_ChrPQArm_GainLoss)==ChromosomeArmSubtypeCNVState_Subset[i,]$ChrPQArm]=ChromosomeArmSubtypeCNVState_Subset[i,]$ChrCNVState
      }
      ChromosomeArmSubtypeCNVState_AllSmp <- dplyr::bind_rows(ChromosomeArmSubtypeCNVState_AllSmp, FinalDF_ChrPQArm_GainLoss)
}

dim(ChromosomeArmSubtypeCNVState_AllSmp) # 47 49

fwrite(ChromosomeArmSubtypeCNVState_AllSmp, file="ChromosomeArmSubtypeCNVState_AllSmp47_v1.2.txt", col.names=TRUE, row.names=FALSE, sep="\t", quote=FALSE )

```

## Step6. Make Supple Figure 2A, bar plot to count the number of Gain/Loss in each chromosome from "ChromosomeArmSubtypeCNVState_AllSmp47.txt" data

```{r}

ChromosomeArmSubtypeCNVGainLoss <- data.table::fread(paste0(dir, "/", ChromosomeArmSubtypeCNVStateFile), header=TRUE, stringsAsFactors=FALSE)
dim(ChromosomeArmSubtypeCNVGainLoss); ChromosomeArmSubtypeCNVGainLoss[1:2,1:5] # 47  50
#      SampleID     Subtype chr01_p chr01_q chr02_p
#        <char>      <char>  <char>  <char>  <char>
# 1:    01-076 G1_MBC_LIKE  

ChrNumber_ColName <- colnames(ChromosomeArmSubtypeCNVGainLoss); print(ChrNumber_ColName)
#  [1] "SampleID"              "Subtype"               "chr01_p"               "chr01_q"               "chr02_p"               "chr02_q"               "chr03_ChromosomeWhole" "chr03_p"              
#  [9] "chr03_q"               "chr04_ChromosomeWhole" "chr04_p"               "chr04_q"               "chr05_p"               "chr05_q"               "chr06_p"               "chr06_q"              
# [17] "chr07_p"               "chr07_q"               "chr08_p"               "chr08_q"               "chr09_p"               "chr09_q"               "chr10_p"               "chr10_q"              
# [25] "chr11_p"               "chr11_q"               "chr12_ChromosomeWhole" "chr12_p"               "chr12_q"               "chr13_q"               "chr14_q"               "chr15_q"              
# [33] "chr16_ChromosomeWhole" "chr16_p"               "chr16_q"               "chr17_p"               "chr17_q"               "chr18_ChromosomeWhole" "chr18_p"               "chr18_q"              
# [41] "chr19_p"               "chr19_q"               "chr20_p"               "chr20_q"               "chr21_p"               "chr21_q"               "chr22_p"               "chr22_q"              
# [49] "NotAvail_NotAvail" 

## Remove 'chr03_p', 'chr04_p', 'chr12_p', 'chr16_ChromosomeWhole', 'chr18_p', 'NotAvail_NotAvail'
ChromosomeArmSubtypeCNVGainLoss_Proc <- ChromosomeArmSubtypeCNVGainLoss %>% dplyr::select(-c(chr03_p, chr04_p, chr12_p, chr16_ChromosomeWhole, chr18_p, NotAvail_NotAvail))
colnames(ChromosomeArmSubtypeCNVGainLoss_Proc)

StackBarplot_Func <- function(ChromosomeArmSubtypeCNVGainLoss=ChromosomeArmSubtypeCNVGainLoss,MySubtype="G1_MBC_LIKE") {
      ChromosomeArmSubtypeCNVGainLoss_MBC <- ChromosomeArmSubtypeCNVGainLoss %>% dplyr::filter(Subtype==MySubtype) %>% dplyr::select(-Subtype) #  %>% 
                                           #  tibble::column_to_rownames("SampleID") %>% data.frame; dim(ChromosomeArmSubtypeCNVGainLoss_MBC) # 10 44
      ChromosomeArmSubtypeCNVGainLoss_MBC[1:2,]
      # SampleID     chr01_p chr01_q chr02_p chr02_q chr03_p chr03_q chr04_p chr04_q chr05_p chr05_q chr06_p chr06_q chr07_p chr07_q chr08_p chr08_q chr09_p chr09_q chr10_p chr10_q chr11_p 
      # 01-076                         
      
      all_chromosome_arms <- ChromosomeArmSubtypeCNVGainLoss_MBC %>% select(starts_with("chr")) %>% colnames()
   
      GainLossCountTable_Long <- ChromosomeArmSubtypeCNVGainLoss_MBC %>%
        # Convert row names into a proper column for pivoting
        #tibble::rownames_to_column(var = "SampleID") %>%
        
        # Pivot the data longer: turn all chromosome columns into two columns (Arm and CNV_Status)
        tidyr::pivot_longer(cols = starts_with("chr"), names_to = "Arm", values_to = "CNV_Status") %>%
        # Filter out all empty/blank strings ("")
        filter(CNV_Status != "") %>%
        # Group by the CNV type (Gain/Loss) and the Chromosome Arm
        group_by(CNV_Status, Arm) %>%
        # Count the number of observations for each combination
        summarise(Count = n(), .groups = 'drop') %>% 
        # 3. Fill missing combinations with Count = 0.
        tidyr::complete( CNV_Status = c("Gain", "Loss"),   Arm=all_chromosome_arms, fill = list(Count = 0) )
      
      print("--- Data Frame for ggplot2 (Long Format) ---")
      print(GainLossCountTable_Long)
      
      # --- 3. ggplot2 Visualization Section ---
      # 3a. Create the stacked bar chart using ggplot2
      cnv_plot_ggplot <- ggplot(GainLossCountTable_Long, aes(x = Arm, y = Count, fill = CNV_Status)) +
        # Use geom_bar with stat="identity" to plot pre-calculated counts
        geom_bar(stat = "identity", position = "stack") +
        # Define custom colors
        scale_fill_manual(values = c("Gain" = "#FF66FF" , "Loss" = "#1b98e0" )) +
        # Add labels and title
        labs( title = "CNV Gain/Loss Counts by Chromosome Arm", x = "Chromosome Arm", y = "Count", fill = "CNV Status" ) + # Legend title
       
        # Apply a clean theme (optional, but highly recommended)
        theme_minimal() +
        # Improve axis text appearance
        theme(
          plot.title = element_text(hjust = 0.5, face = "bold"), # Center title
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels
          ### Remove grid lines
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
        )
      
      PlotHeight=max(GainLossCountTable_Long$Count)*0.4
      if(MySubtype=="G1_MBC_LIKE" ) PlotHeight=2.5
      if(MySubtype=="G2_PC_LIKE" ) PlotHeight=4
      ggsave(filename = paste0(dir , "/", paste0("SuppleFig2A_StackBarplot_CountGainLossPerChr_",gsub("G1_|G2_|G3_|_","",MySubtype),"v1.2.pdf")) , plot = cnv_plot_ggplot,  width=12, height=PlotHeight,  units = "in")
} 

### 1) MBC-like subset
StackBarplot_Func(ChromosomeArmSubtypeCNVGainLoss=ChromosomeArmSubtypeCNVGainLoss_Proc, MySubtype="G1_MBC_LIKE")

### 2) PC-like subset
StackBarplot_Func(ChromosomeArmSubtypeCNVGainLoss=ChromosomeArmSubtypeCNVGainLoss_Proc, MySubtype="G2_PC_LIKE")

### 3) NotAvail subset
StackBarplot_Func(ChromosomeArmSubtypeCNVGainLoss=ChromosomeArmSubtypeCNVGainLoss_Proc, MySubtype="G3_NotAvail")


```

## Step 7. GenVisR plot and Segment horizontal line for Chr6q Region for all 47s, MBC-like, PC-like or NotAvail

Extract only chr6q deletions and see the segment size by MBC-like and PC-like. Use cnFreq_Modified() use at "Source_GenVisR_Modified_ByCNVSegment_BycnQual.R"

```{r}
### 1) all Chr6q regions for All 47 smp
table(cnData_NoWM91WM93WM94$sample); length(table(cnData_NoWM91WM93WM94$sample)); dim(cnData_NoWM91WM93WM94) # 47  # 2142 17

cnData_Chr6qOnly <- cnData_NoWM91WM93WM94 %>% dplyr::filter(chromosome=="chr6"); dim(cnData_Chr6qOnly) # 69 17
table(cnData_Chr6qOnly$sample); length(table(cnData_Chr6qOnly$sample)) # 24  
saveRDS(cnData_Chr6qOnly, "cnData_Chr6qOnly.rds")
## dplyr::filter(chromosome=="chr6", start > 61*10^6 ) # 32 17 <<<<<<<<==========This exclude samples in the group that don't have loss in Chr6q
## so GenVisR plot couldn't reflect the exact number of samples. So, I added one line in "Source_GenVisR_Modified_ByCNVSegments_BycnQual.R

CentromereRegion <- ggplot2::geom_vline(xintercept = c(59800000, 59800001))
CNVPlot_Chr6DelRegion <- cnFreq_Modified(x=cnData_Chr6qOnly, genome="hg38", plotChr = "chr6", plotLayer = CentromereRegion, WMSubtype="AllType", ChrNumb="Chr6q", 
                                         ChrStartPosition=59800000, ChrEndPosition=170805979)
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr6qOnly_All24smp_v1.6_MoreXaxisLabel.pdf", CNVPlot_Chr6DelRegion, width=8, height=4)

### 2) Chr6q - MBC-LIKE
cnData_NoWM91WM93WM94_MBCLike_Chr6q <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr6" ) %>% dplyr::filter(Subtype=="G1_MBC_LIKE") %>% select(-Subtype) ; dim(cnData_NoWM91WM93WM94_MBCLike_Chr6q) 
## dplyr::filter(chromosome=="chr6", start > 61*10^6 ) # 32 17 <<<<<<<<==========This exclude samples in the group that don't have loss in Chr6q
## so GenVisR plot couldn't reflect the exact number of samples. So, I added one line in "Source_GenVisR_Modified_ByCNVSegments_BycnQual.R

saveRDS(cnData_NoWM91WM93WM94_MBCLike_Chr6q, "cnData_NoWM91WM93WM94_MBCLike_Chr6q.rds")

table(cnData_NoWM91WM93WM94_MBCLike_Chr6q$sample)
# 01-076 01-190   WM31   WM33   WM46   WM47   WM52   WM65   WM70    WM9 
#      1     29      1      3      1      3      1      1      2      1 

CentromereRegion <- ggplot2::geom_vline(xintercept = c(59800000, 59800001))
CNVPlot_Chr6qDelRegion_MBC <- cnFreq_Modified(x=cnData_NoWM91WM93WM94_MBCLike_Chr6q, genome="hg38", plotChr="chr6", plotLayer=CentromereRegion,  WMSubtype="MBCLike", ChrNumb="Chr6q", 
                                              ChrStartPosition=59800000, ChrEndPosition=170805979)
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr6qOnly_MBC3smp_v1.6_MoreXaxisLabel.pdf", CNVPlot_Chr6qDelRegion_MBC, width=6, height=4)

### 3) Chr6q - PC-LIKE
cnData_NoWM91WM93WM94_PCLike_Chr6q <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr6") %>% dplyr::filter(Subtype=="G2_PC_LIKE") %>% select(-Subtype) ; dim(cnData_NoWM91WM93WM94_PCLike_Chr6q) # 52 17
## dplyr::filter(chromosome=="chr6", start > 61*10^6 ) %>%

saveRDS(cnData_NoWM91WM93WM94_PCLike_Chr6q, "cnData_NoWM91WM93WM94_PCLike_Chr6q.rds")

table(cnData_NoWM91WM93WM94_PCLike_Chr6q$sample); length(table(cnData_NoWM91WM93WM94_PCLike_Chr6q$sample)) # 12 
# 01-115 01-131 01-163 04-006   WM22   WM26   WM27   WM29   WM30   WM32   WM37   WM42   WM43   WM48   WM49   WM54   WM66   WM68   WM69   WM72   WM73    WM8 
#      3      3      1      1      3      1      1      3      3      2      3      6      3      3      1      1      2      2      3      1      2      4 
     
CNVPlot_Chr6qDelRegion_PC_Chr6q <- cnFreq_Modified(x=cnData_NoWM91WM93WM94_PCLike_Chr6q, genome="hg38", plotChr="chr6", plotLayer=CentromereRegion,  WMSubtype="PCLike", ChrNumb="Chr6q", 
                                                     ChrStartPosition=59800000, ChrEndPosition=170805979) 
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr6qOnly_PC12smp_v1.6_MoreXaxisLabel.pdf", CNVPlot_Chr6qDelRegion_PC_Chr6q, width=6, height=4)


### 4) Chr6q - NotAvail
cnData_NoWM91WM93WM94_NotAvail_Chr6q <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr6") %>% 
                                  dplyr::filter(Subtype=="G3_NotAvail") %>% select(-Subtype);  # , start > 61*10^6 
dim(cnData_NoWM91WM93WM94_NotAvail_Chr6q) # 45 17

saveRDS(cnData_NoWM91WM93WM94_NotAvail_Chr6q, "cnData_NoWM91WM93WM94_NotAvail_Chr6q.rds")

# table(cnData_NoWM91WM93WM94_NotAvail_Chr6q$sample); length(table(cnData_NoWM91WM93WM94_NotAvail_Chr6q$sample)) # 15
# WM67 WM74 WM77 WM80 WM82 WM83 WM84 WM85 WM86 WM87 WM89 WM90 WM95 WM96 WM97 
#    3    1    1    1   10    1    4    3    4    5    2    3    1    5    1 

CNVPlot_Chr6qDelRegion_NotAvail_Chr6q <- cnFreq_Modified(x=cnData_NoWM91WM93WM94_NotAvail_Chr6q, genome="hg38", plotChr="chr6", plotLayer=CentromereRegion,  WMSubtype="NotAvail", ChrNumb="Chr6q", 
                                                          ChrStartPosition=59800000, ChrEndPosition=170805979) 
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr6qOnly_NotAvail5smp_v1.6_MoreXaxisLabel.pdf", CNVPlot_Chr6qDelRegion_NotAvail_Chr6q, width=6, height=4)


```

#### Extract genes in a certain region of Chromosome 6.

```{r}
Chr6GeneAnnot <- fread(Chr6GeneAnnotFile, header=FALSE, stringsAsFactors=FALSE); dim(Chr6GeneAnnot); Chr6GeneAnnot[1:2,] #  95165     9
colnames(Chr6GeneAnnot) <- c("ChrNumb","Curated","ExonIntrol","Start","End","Score","Strand","Sign", "Info"); Chr6GeneAnnot[1:2,]

Chr6GeneAnnot_Gene <- Chr6GeneAnnot %>% dplyr::filter(ExonIntrol=="gene"); dim(Chr6GeneAnnot_Gene) #  2529    9

MSB_Start <- 99637055;  # StartX[1] in MSB-Like
MSB_End <- 99654926     # EndX[1] in MSB-Like  ## 0 genes

MSB_Start <-155930053   # StartX[27] in MSB-Like
MSB_End <- 170741611     # EndX[27] in MSB-Like  ## 290 genes. Black line. 

MSB_Start <- 79967200;  # StartX[2] in MSB-Like
MSB_End <- 80080718     # EndX[2] in MSB-Like  ## 3 

MSB_Start <- 140759906;  # StartX[3] in MSB-Like
MSB_End <- 140918400     # EndX[3] in MSB-Like  ## # same as StartX[1] and EndX[1]

MSB_Start <- 90929121;  # StartX[6] in MSB-Like
MSB_End <- 91166300     # EndX[6] in MSB-Like  ## # 0 

MSB_Start <- 91167400;  # StartX[7] in MSB-Like
MSB_End <- 91416000     # EndX[7] in MSB-Like  ## # 1

MSB_Start <-73536500   # StartX[26] in MSB-Like
MSB_End <- 79965809     # EndX[26] in MSB-Like # 78 


Chr6GeneAnnot_GeneInRegion <- Chr6GeneAnnot_Gene %>% dplyr::filter( (Start > MSB_Start) & (End < MSB_End) ); dim(Chr6GeneAnnot_GeneInRegion) # 169 9


```

## GenVisR plot and Segment horizontal line for Chr1q Region for all 47s, MBC-like, PC-like or NotAvail \<\<\<\<\<\<=== Chr1q

Extract onley chr1q deletions and see the segment size by MBC-like and PC-like. Use cnFreq_Modified() at "Source_GenVisR_Modified_ByCNVSegment_BycnQual.R"

```{r}
### ============== 1) all Chr1q regions for All 47 smp
table(cnData_NoWM91WM93WM94$sample); length(table(cnData_NoWM91WM93WM94$sample)) # 47 

cnData_Chr1qOnly <- cnData_NoWM91WM93WM94 %>% dplyr::filter(chromosome=="chr1"); dim(cnData_Chr1qOnly) # 66 17 # dplyr::filter(chromosome=="chr1", start > 1.217*10^8 )
table(cnData_Chr1qOnly$sample); length(table(cnData_Chr1qOnly$sample)) # 47

CentromereRegion <- ggplot2::geom_vline(xintercept = c(123400000, 123400001))
CNVPlot_Chr1DelRegion <- cnFreq_Modified(x=cnData_Chr1qOnly, genome="hg38", plotChr="chr1", plotLayer=CentromereRegion, WMSubtype="AllType", ChrNumb="Chr1q",
                                       ChrStartPosition=123300000, 	ChrEndPosition=248956422 )
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr1qOnly_All30smp_v1.6_MoreXaxisLabel.pdf", CNVPlot_Chr1DelRegion, width=8, height=4)

### ============== 2) Chr1q - MBC-LIKE
cnData_Chr1q_NoWM91WM93WM94_MBCLike <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr1", start > 1.217*10^8 ) %>% dplyr::filter(Subtype=="G1_MBC_LIKE") %>% select(-Subtype) ; dim(cnData_Chr1q_NoWM91WM93WM94_MBCLike) # 15 17

saveRDS(cnData_Chr1q_NoWM91WM93WM94_MBCLike, "cnData_Chr1q_NoWM91WM93WM94_MBCLike.rds")

table(cnData_Chr1q_NoWM91WM93WM94_MBCLike$sample)  ## 6 subjects
# 01-190   WM31   WM33   WM52   WM65   WM70 
#      8      1      2      1      2      1 

CentromereRegion <- ggplot2::geom_vline(xintercept = c(123400000, 123400001))
CNVPlot_Chr1DelRegion_MBC <- cnFreq_Modified(x=cnData_Chr1q_NoWM91WM93WM94_MBCLike, genome="hg38", plotChr="chr1", plotLayer=CentromereRegion,  WMSubtype="MBCLike",  ChrNumb="Chr1q", 
                                               ChrStartPosition=123300000, 	ChrEndPosition=248956422 )
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr1qOnly_MBC6smp_v1.5_MoreXaxisLabel.pdf", CNVPlot_Chr1DelRegion_MBC, width=6, height=4)

### ============== 3) Chr1q - PC-LIKE
cnData_Chr1q_NoWM91WM93WM94_PCLike <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr1", start > 1.217*10^8 ) %>% dplyr::filter(Subtype=="G2_PC_LIKE") %>% select(-Subtype) ; dim(cnData_Chr1q_NoWM91WM93WM94_PCLike) # 29  17

table(cnData_Chr1q_NoWM91WM93WM94_PCLike$sample); length(table(cnData_Chr1q_NoWM91WM93WM94_PCLike$sample)) # 16
#   01-115 01-131   WM22   WM26   WM27   WM29   WM30   WM32   WM43   WM49   WM66   WM68   WM69   WM72   WM73    WM8 
#    3      3      1      2      2      1      1      2      2      2      2      1      2      2      2      1 

CNVPlot_Chr1DelRegion_PC <- cnFreq_Modified(x=cnData_Chr1q_NoWM91WM93WM94_PCLike, genome="hg38", plotChr="chr1", plotLayer=CentromereRegion,  WMSubtype="PCLike", ChrNumb="Chr1q",
                                              ChrStartPosition=123300000, 	ChrEndPosition=248956422 ) 
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr1qOnly_PC16smp_v1.5_MoreXaxisLabel.pdf", CNVPlot_Chr1DelRegion_PC, width=6, height=4)

### ============== 4) Chr1q - NotAvail
cnData_Chr1q_NoWM91WM93WM94_NotAvail <- cnData_NoWM91WM93WM94_Metadata %>% dplyr::filter(chromosome=="chr1", start >1.217*10^8  ) %>% dplyr::filter(Subtype=="G3_NotAvail") %>% select(-Subtype);
dim(cnData_Chr1q_NoWM91WM93WM94_NotAvail) # 22 17

table(cnData_Chr1q_NoWM91WM93WM94_NotAvail$sample); length(table(cnData_Chr1q_NoWM91WM93WM94_NotAvail$sample)) # 8
#   WM74 WM77 WM83 WM84 WM86 WM87 WM90 WM95 
#    2    1    4    2    2    2    1    8 

CNVPlot_Chr1DelRegion_NotAvail <- cnFreq_Modified(x=cnData_Chr1q_NoWM91WM93WM94_NotAvail, genome="hg38", plotChr="chr1", plotLayer=CentromereRegion,  WMSubtype="NotAvail",  ChrNumb="Chr1q", 
                                                    ChrStartPosition=123300000, 	ChrEndPosition=248956422 ) 
ggplot2::ggsave(filename="GenVisR_CNVPlot_Chr1qOnly_NotAvail8smp_v1.5_MoreXaxisLabel.pdf", CNVPlot_Chr1DelRegion_NotAvail, width=6, height=4)


```
